A Recurrence

### 2014-12-08 draft.tex
Section 4 done

In this section we consider the case where $F(x,y) = f(y)/x$ where $f(y) = y^n$, where $n$ can be any complex number, not just an integer. The question we would like to answer is given an integer $p$ can we always come up with an $n$ such that $F$ has period $p$. We first explore some basic properties of the sequences such functions produce and compute an explicit example. We then give an explicit formula for the $k$th term in the sequence. Finally using a generating function we are able to produce a necessary condition for $n$, to produce a sequence with period $p$. Using Mathematica we verified numerically for values of $p$ up to 1000 that this necessary condition is also sufficient and because of this we conjecture 
\begin{conjecture}
\label{cos}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}

We begin by considering the sequence produced by $F(x,y) = y^n/x$ 
\begin{equation*}
x^1y^0, x^0y^1, x^{-1}y^{n}, x^{-n}y^{n^2-1}, x^{-n^2+1}y^{n^3-2n},x^{-n^3+n^2+n}y^{n^4-3n^2+1}, \dots \,.
\end{equation*}

The interesting thing to note is that all of the terms in the sequence are now monomials, which will simplify our analysis. Now let $\alpha_k$ be the sequence of the exponents of $x$ in the above sequence, and let $\beta_k$ be the sequence of exponents of $y$. We can easily compute the first few terms of the sequence
\begin{align*}
\alpha_0 &= 1  &\beta_0& = 0\\
\alpha_1 &= 0 &\beta_1& = 1\\
\alpha_2 &= -1  &\beta_2& = n\\
\alpha_3 &= -n &\beta_3& = n^2-1\\
\alpha_4 &= -n^2+1 &\beta_4& = n^3-2n\\
\alpha_5 &= -n^3+2n &\beta_5& =n^4-3n^2+1\\ 
\alpha_6 &= -n^4+3n^2 -1 &\beta_6& =n^5-4n^3+3n \\ 
\alpha_7 &= -n^5+4n^3 - 3n &\beta_7& =n^6-5n^4+6n^2 - 1.
\end{align*}

We can use these $\{\alpha_i\}$ and $\{\beta_i\}$ sequences to get values of $n$ that achieve certain periods. For example, if we want a function of the form $F(x,y) = y^n/x$ with period 5, we simply set $\alpha_5 = \alpha_0 = 1$, $\beta_5 = \beta_0 = 0$, $\alpha_6 = \alpha_1 = 0$, and $\beta_6 = \beta_1 = 1$. This gives us a system of four equations in one variable
\begin{align*}
-n^3+2n &= 1\\
n^4-3n^2+1&=0\\
-n^4+3n^2 -1&= 0\\
n^5-4n^3+3&n=1\\
\end{align*}
There is no guarantee that this system has a solution, but it turns out that setting $n = -\frac{1+\sqrt{5}}{2} = \phi$, the golden ratio, works. This tells us that $F(x,y) = y^{-\phi}/x$ is periodic with period 5, and we can indeed check that this is the case.

We now give an explicit form for $\alpha_k$ and $\beta_k$ in terms of $k$ and $n$. Unfortunately we are not able to show that the equations these produce always have common solutions. 
We first notice that $\alpha_k$ and $\beta_k$ are similar sequences, as they can be defined as
\begin{align*}
\alpha_0 = 1, \alpha_1 = 0, \alpha_k = n\alpha_{k-1} - \alpha_{k-2}\\
\beta_0 = 0, \beta_1 = 1, \beta_k = n\beta_{k-1} - \beta_{k-2}
\end{align*}

@@

We note further that there is a relation between different terms of the two sequences.  
\begin{lemma} 
\label{ab} For $k\geq 0$
\begin{equation*}
\alpha_{k+1} = - \beta_{k}.
\end{equation*}
\end{lemma}
\begin{proof}
We will show this using induction. As a base case note that $\alpha_1 = 0 = \beta_0$, and $\alpha_2 = -n = \beta_1$. Now if we assume this holds for $k \leq m$ and would like to show it holds for $k = m+1$ consider $\alpha_{m+1}$
\begin{equation*}
\alpha_{m+1} = n \alpha_{m} - \alpha_{m-1}.
\end{equation*}
Well $\alpha_m = -\beta_{m-1}$ and $\alpha_{m-1} = - \beta_{m-2}$ and so 
\begin{equation*}
\alpha_{m+1} = - n \beta_{m-1} + \beta_{m-2} = -\beta_{m},
\end{equation*}
as desired.
\end{proof}

\begin{proposition} 
For $k \geq 1$
\begin{align*}
\beta_k = \binom{k-1}{k-1} n^{k-1} - \binom{k-2}{k-3}n^{k-3} +\binom{k-3}{k-5}n^{k-5} -\binom{k-4}{k-7}n^{k-7} + \dots \\
=\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i}\\
\end{align*}
\end{proposition}
\begin{proof}
It is easy to check that this holds for $k = 1$ and $k = 2$.  If we plug it in to the recurrence relation
\begin{equation*}
\beta_k = n\beta_{k-1} - \beta_{k-2},
\end{equation*}
we must have 
\begin{align*}
\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i} = & n \sum_{i = 0}^{\lfloor (k-2)/2 \rfloor} (-1)^i\binom{k-2-i}{k-2-2i}n^{k-2-2i} \\
& +\sum_{i = 0}^{\lfloor (k-3)/2 \rfloor} (-1)^i\binom{k-3-i}{k-3-2i}n^{k-3-2i}\\
\end{align*}
Checking this equation when $k$ is even and when $k$ is odd, as well as using combinatorial identities, shows that this expression for $\beta_k$ satisfies the recurrence relation. Because of this the expression must equal $\beta_k$ for $k \geq 1$.
\end{proof}

So we have an expression for $\beta_k$, which in turn gives us an expression for $\alpha_k$ by Lemma \ref{ab}. Thus, we know that if we want to attain a specific period $p$, we have to solve four equations involving these known expressions.

In fact thes four equations are 
\begin{align*}
\alpha_p &=1 \\
\beta_p &=0\\ 
\alpha_{p+1} &= 0 \\
\beta_{p+1} &= 1.
\end{align*}
As shown in Lemma \ref{ab} above $\alpha_{p+1} = -\beta_p$,  $\beta_p=0$ and $\alpha_{p+1} =0$ are the same equation so we actually only need to find a value for $n$ which satisfies the three equations
\begin{align*}
\alpha_p &=1 \\
\beta_p &=0\\ 
\beta_{p+1} &= 1.
\end{align*}
It remains to show that these equations always produce a real solution for $n$.
\\

Now using a generating function approach we give a necessary condition for $n$. 
First we must produce the generating function for $\alpha_k$. Recall that a generating function of an infinite sequence is a formal power series with coefficients equal to the terms of the sequence.  So if $G(\xi)$ is the generating function of $\alpha_k$, then 
\begin{equation*}
G(\xi) = \sum_{k=0}^{\infty} \alpha_k \xi^k,
\end{equation*}
where $\xi$ is some formal variable. We can find a closed form expression for $G(\xi)$.

\begin{lemma} For a given $n$ the generating function for $\alpha_k$ is 
\begin{equation*}
G(\xi) = \frac{ 1 - n\xi}{1 - n \xi + \xi^2} .
\end{equation*}
\end{lemma}
\begin{proof}
Recall the recurrence relation for $\alpha_k$ 
\begin{equation*}
\alpha_k = n \alpha_{k-1} - \alpha_{k-2}, 
\end{equation*}
and note that 
\begin{align*}
G(\xi) &= \alpha_0 + &\alpha_1 \xi + &\alpha_2 \xi^2 + \cdots \\
n \xi G(\xi) &= &n\alpha_0 \xi + &n\alpha_1 \xi^2 + n\alpha_2 \xi^3 + \cdots \\
\xi^2 G(\xi) &= &&\alpha_0 \xi^2 + \alpha_1 \xi^3 + \alpha_2 \xi^4 + \cdots.
\end{align*}
So if we subtract the second and third equations from the first and simplify using the recurrence relation we see 
\begin{equation*}
G(\xi) - n \xi G(\xi) - \xi^2 G(\xi) = \alpha_0 + \alpha_1 \xi - n \alpha \xi.
\end{equation*}
Rearranging this equation we can see 
\begin{equation*}
G( \xi) (1-n \xi + \xi^2) = n\alpha_0 + \alpha_1 \xi - n\alpha_0 \xi.
\end{equation*}
Using the fact that $\alpha_0=1, \alpha_1=0$ and rearranging gives us
\begin{equation*}
G(\xi) = \frac{ 1 - n\xi}{1 - n \xi + \xi^2},
\end{equation*}
as desired.
\end{proof}

We can also find another expression for $G(\xi)$ if we assume that the sequence is periodic. 
\begin{lemma} If $n$ is such that $F$ produces a sequence of period $p$ then the generating function for $\alpha_k$ is 
\begin{equation*}
G(\xi) = \frac{1}{1- \xi^p} (\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}).
\end{equation*}
\end{lemma}
\begin{proof}
In order for the period of $F$ to be $p$ we must have $\alpha_p= \alpha_0$ and $\alpha_{p+1} = \alpha_1$. Plugging that into the power series form of $G(\xi)$ tells us that
\begin{align*}
G(\xi) = &\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1} \\
+ \xi^p(&\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}) \\ 
+ \cdots \, ,
\end{align*}
and so using the summation formula for geometric series we have
\begin{equation*}
G(\xi) = \frac{1}{1- \xi^p} (\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}).
\end{equation*}
\end{proof}

Using these two expressions for the generating function we can produce the following necessary condition for $n$. 
\begin{proposition} 
\label{must}
If $n$ is such that $F$ produces a sequence of period $p$ then $n$ must be of the form
\begin{equation*}
n= 2 \cos(2 \pi l/p), 
\end{equation*}
for some $l \in \{0,1,2, \ldots, p-1\}$. 
\end{proposition}
\begin{proof}
If we equate the two different expressions for $G(\xi)$ we have 
\begin{equation*}
\frac{ 1 - n\xi}{1 - n \xi + \xi^2} = \frac{1}{1- \xi^p} (\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}).
\end{equation*}

@@

Note also that the locations of discontinuities for both functions must be the same. Therefore we must have that the zeroes of $1-n\xi + \xi^2$ are the same as those zeroes of the denominator of the left hand side which were not canceled out. Thus for some $l$ and $m$, we must have 
\begin{align*}
(\xi - e^{2 \pi i l/p})(\xi - e^{2 \pi i m/p}) = 1 - n \xi + \xi^2,
\end{align*}
or 
\begin{align*}
n = e^{2 \pi i l/p} + e^{2 \pi i m/p} \\
1 = e^{2 \pi i (l+m)/p}.
\end{align*}
So $n$ must be the sum of some $p$th roots of unity, and in fact these roots of unity must be conjugates of one another, so $n$ is always real. Therefore  
\begin{equation*}
n= 2 \cos(2 \pi l/p), 
\end{equation*}
for some $l \in \{0, 1,2, \ldots, p-1\}$. 
\end{proof}
Note that by Proposition \ref{period6} we know that $F(x,y)=y/x$ has period 6, and since a function $F$ can only have one period for $p \neq 6$, $l$ cannot equal 0.

@@

\begin{conjecture} For all $p$ there exists $n$ such that $F(x,y)=y^n/x$ has period $p$. 
\end{conjecture} 
This conjecture is equivalent to saying that we can always find a value of $n$ that satisfies the three equations 
\begin{align*}
\alpha_p &=1\\ 
\beta_p &=0\\
\beta_{p+1} &=1.
\end{align*}
We have verified that solutions exist for values of $p$ up to 1000. Note that this conjecture along with Proposition \ref{must} are not equivalent to Conjecture \ref{cos}, which for ease of reading we restate here. 
\begin{conjecture}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}
Conjecture \ref{cos} states that for \emph{any} value of $l$ the $n$ produced has the desired period, while Conjecture

>>>>

In this section we consider the case where $F(x,y) = f(y)/x$ with $f(y) = y^n$, where $n$ is any complex number, not just an integer. The question we would like to answer is given an integer $p$ can we always come up with an $n$ such that $F$ has period $p$. We first explore some basic properties of the sequences such functions produce and compute an explicit example. We then give an explicit formula for the $k$th term in the sequence. Finally using a generating function we are able to produce a necessary condition on $n$, in order for $F$ to produce a sequence with period $p$. Using Mathematica we were able to verify numerically for values of $p$ up to 1000 that this necessary condition is also sufficient. Because of this we conjecture 
\begin{conjecture}
\label{cos}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}

We begin by considering the sequence produced by $F(x,y) = y^n/x$ 
\begin{equation}
\label{exponents}
x^1y^0, x^0y^1, x^{-1}y^{n}, x^{-n}y^{n^2-1}, x^{-n^2+1}y^{n^3-2n},x^{-n^3+n^2+n}y^{n^4-3n^2+1}, \dots \,.
\end{equation}

We note that all of the terms in the sequence are now monomials, which simplifies our analysis. For simplicity we introduce the following notation.
\begin{definition} 
Let $\alpha_k$ be the sequence of exponents of $x$ in Equation \ref{exponents}, and $\beta_k$ be the sequence of exponents of $y$ in Equation \ref{exponents}.
\end{definition}
We can easily compute the first few terms of the sequence
\begin{align*}
\alpha_0 &= 1  &\beta_0& = 0\\
\alpha_1 &= 0 &\beta_1& = 1\\
\alpha_2 &= -1  &\beta_2& = n\\
\alpha_3 &= -n &\beta_3& = n^2-1\\
\alpha_4 &= -n^2+1 &\beta_4& = n^3-2n\\
\alpha_5 &= -n^3+2n &\beta_5& =n^4-3n^2+1\\ 
\alpha_6 &= -n^4+3n^2 -1 &\beta_6& =n^5-4n^3+3n \\ 
\alpha_7 &= -n^5+4n^3 - 3n &\beta_7& =n^6-5n^4+6n^2 - 1.
\end{align*}
We can use these sequences $\{\alpha_i\}$ and $\{\beta_i\}$ to compute values for $n$ that achieve certain periods. In particular if we can find an $n$ which solves the four equations 
\begin{align*}
\alpha_p &= 1 \\
\beta_p &=0\\
\alpha_{p+1} &=  0\\
\beta_{p+1}&=1,
\end{align*}
then $F(x,y)=y^n/x$ has period $p$.

For example, if we want a function of the form $F(x,y) = y^n/x$ with period 5, we simply set $\alpha_5 = \alpha_0 = 1$, $\beta_5 = \beta_0 = 0$, $\alpha_6 = \alpha_1 = 0$, and $\beta_6 = \beta_1 = 1$. This gives us a system of four equations in one variable
\begin{align*}
-n^3+2n &= 1\\
n^4-3n^2+1&=0\\
-n^4+3n^2 -1&= 0\\
n^5-4n^3+3&n=1\\
\end{align*}
There is no guarantee that this system has a solution, but in fact $n = -\frac{1+\sqrt{5}}{2} = -\phi$, the golden ratio, is a solution. This tells us that $F(x,y) = y^{-\phi}/x$ is periodic with period 5, and we can check that indeed this is the case.

We now give explicit formulas for $\alpha_k$ and $\beta_k$ in terms of $k$ and $n$. Unfortunately we are not able to show that the equations these produce always have common solutions. 
We first notice that $\alpha_k$ and $\beta_k$ are similar sequences, as they can be defined as
\begin{align*}
\alpha_0 = 1, \alpha_1 = 0, \alpha_k = n\alpha_{k-1} - \alpha_{k-2}\\
\beta_0 = 0, \beta_1 = 1, \beta_k = n\beta_{k-1} - \beta_{k-2}
\end{align*}

@@

We note further that there is a relation between different terms of the two sequences.  
\begin{lemma} 
\label{ab} For $k\geq 0$
\begin{equation*}
\alpha_{k+1} = - \beta_{k}.
\end{equation*}
\end{lemma}
\begin{proof}
We will show this using induction. As a base case note that $\alpha_1 = 0 = \beta_0$, and $\alpha_2 = -n = \beta_1$. Now if we assume this holds for $k \leq m$ and would like to show it holds for $k = m+1$ consider $\alpha_{m+1}$ which by the recurrence relation is
\begin{equation*}
\alpha_{m+1} = n \alpha_{m} - \alpha_{m-1}.
\end{equation*}
Well $\alpha_m = -\beta_{m-1}$ and $\alpha_{m-1} = - \beta_{m-2}$ by the induction hypothesis and so 
\begin{equation*}
\alpha_{m+1} = - n \beta_{m-1} + \beta_{m-2} = -\beta_{m}.
\end{equation*}
\end{proof}
Note that this Lemma also reduces the number of simultaneous equations that we need to solve from four to three as we always have $\alpha_{p+1} = - \beta_{p}$. Both are set to zero so it is enough to only include one of them in the system of equations

We now given an exact formula for $\beta_k$ 
\begin{proposition} 
For $k \geq 1$
%= \binom{k-1}{k-1} n^{k-1} - \binom{k-2}{k-3}n^{k-3} +\binom{k-3}{k-5}n^{k-5} -\binom{k-4}{k-7}n^{k-7} + \dots \\
\begin{align*}
\beta_k =\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i}
\end{align*}
\end{proposition}
\begin{proof}
It is easy to check that this holds for $k = 1$ and $k = 2$.  If we plug the formula into the recurrence relation
\begin{equation*}
\beta_k = n\beta_{k-1} - \beta_{k-2},
\end{equation*}
we must have 
\begin{align*}
\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i} = & n \sum_{i = 0}^{\lfloor (k-2)/2 \rfloor} (-1)^i\binom{k-2-i}{k-2-2i}n^{k-2-2i} \\
& +\sum_{i = 0}^{\lfloor (k-3)/2 \rfloor} (-1)^i\binom{k-3-i}{k-3-2i}n^{k-3-2i}\\
\end{align*}
Checking this equation separately when $k$ is even or odd, as well as using combinatorial identities, shows that this expression for $\beta_k$ satisfies the recurrence relation. Because of this our formula is valid for $k \geq 1$.
\end{proof}

This expression for $\beta_k$, gives us an expression for $\alpha_k$ by Lemma \ref{ab}. In order to obtain a function $F$ with period $p$ we must solve the three equations
\begin{align*}
\alpha_p &=1 \\
\beta_p &=0\\ 
\beta_{p+1} &= 1,
\end{align*}
plugging in the explicit values for $\alpha_p, \beta_p$ and $\beta_{p+1}$. We were not able to show that these equations always produce a solution for $n$.
\\

Now using a generating function approach, given a desired period $p$, we give a necessary condition for $n$. 
First we produce the generating function for $\alpha_k$. A generating function of an infinite sequence is a formal power series with coefficients equal to the terms of the sequence.  So if $G(\xi)$ is the generating function of $\alpha_k$, then 
\begin{equation*}
G(\xi) = \sum_{k=0}^{\infty} \alpha_k \xi^k,
\end{equation*}
where $\xi$ is some formal variable. We can find a closed form expression for $G(\xi)$.

\begin{lemma} The generating function for $\alpha_k$ is 
\begin{equation*}
G(\xi) = \frac{ 1 - n\xi}{1 - n \xi + \xi^2} .
\end{equation*}
\end{lemma}
\begin{proof}
Recall the recurrence relation for $\alpha_k$ 
\begin{equation*}
\alpha_k = n \alpha_{k-1} - \alpha_{k-2}, 
\end{equation*}
and note that 
\begin{align*}
G(\xi) &= \alpha_0 + \> \alpha_1 \xi + \>\> \alpha_2 \xi^2 + \cdots \\
n \xi G(\xi) &= \qquad n\alpha_0 \xi + n\alpha_1 \xi^2 + n\alpha_2 \xi^3 + \cdots \\
\xi^2 G(\xi) &= \qquad \qquad \quad \quad \alpha_0 \xi^2 + \alpha_1 \xi^3 +  \cdots.
\end{align*}
So if we subtract the second and third equations from the first, simplify using the recurrence relation and rearrange terms we obtain
\begin{equation*}
G( \xi) (1-n \xi + \xi^2) = \alpha_0 + \alpha_1 \xi - n\alpha_0 \xi.
\end{equation*}
Using the fact that $\alpha_0=1, \alpha_1=0$ and rearranging gives the desired result.
\end{proof}

We can also find another expression for $G(\xi)$ if we assume that the sequence is periodic. 
\begin{lemma} If $n$ is such that $F$ produces a sequence of period $p$ then the generating function for $\alpha_k$ is 
\begin{equation*}
G(\xi) = \frac{1}{1- \xi^p} (\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}).
\end{equation*}
\end{lemma}
\begin{proof}
In order for the period of $F$ to be $p$ we must have $\alpha_p= \alpha_0$ and $\alpha_{p+1} = \alpha_1$. Plugging that into the power series form of $G(\xi)$ tells us that
\begin{align*}
G(\xi) =&  \qquad \> \alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1} \\
 &+ \xi^p(\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}) \\ 
 &+ \xi^{2p}(\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}) \\ 
 &+ \cdots \, ,
\end{align*}
and using the summation formula for geometric series gives us our desired result.
\end{proof}

Using these two expressions for the generating function we can produce the following necessary condition for $n$. 
\begin{proposition} 
\label{must}
If $F$ produces a sequence of period $p$ then $n$ must be of the form
\begin{equation*}
n= 2 \cos(2 \pi l/p), 
\end{equation*}
for some $l \in \{1,2, \ldots, p-1\}$. 
\end{proposition}
\begin{proof}
Both expressions for $G(\xi)$ are valid and so we can equate them 
\begin{equation*}
\frac{ 1 - n\xi}{1 - n \xi + \xi^2} = \frac{1}{1- \xi^p} (\alpha_0 + \alpha_1 \xi + \alpha_2 \xi^2 + \cdots + \alpha_{p-1} \xi^{p-1}).
\end{equation*}

@@

Note also that the locations of discontinuities for both functions must be the same. Therefore the zeroes of $1-n\xi + \xi^2$ must be the same as those zeroes of the denominator of the right hand side which were not canceled out. Thus for some $s$ and $t$, we must have 
\begin{align*}
(\xi - e^{2 \pi i s/p})(\xi - e^{2 \pi i t/p}) = 1 - n \xi + \xi^2,
\end{align*}
or written another way
\begin{align*}
n = e^{2 \pi i s/p} + e^{2 \pi i t/p} \\
1 = e^{2 \pi i (s+t)/p}.
\end{align*}
So $n$ must be the sum of two distinct $p$th roots of unity, and in fact these roots of unity must be conjugates of one another, so $n$ is always real. Therefore  
\begin{equation*}
n= 2 \cos(2 \pi l/p), 
\end{equation*}
for some $l \in \{1,2, \ldots, p-1\}$. 
\end{proof}

@@

\begin{conjecture}
\label{exists}
 For all $p$ there exists $n$ such that $F(x,y)=y^n/x$ has period $p$. 
\end{conjecture} 
This conjecture is equivalent to saying that we can always find a value of $n$ that satisfies the three equations 
\begin{align*}
\alpha_p &=1\\ 
\beta_p &=0\\
\beta_{p+1} &=1.
\end{align*}
We have been able to verify that solutions exist for values of $p$ up to 1000. Note that this conjecture along with Proposition \ref{must} are not equivalent to Conjecture \ref{cos}, which for ease of reading we restate here. 
\begin{conjecture}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}
Conjecture \ref{cos} states that for \emph{any} value of $l$ the $n$ produced has the desired period, while Conjecture \ref{exists} along with Proposition \ref{must} states that one such $l$ exists such that the $n$ produced has the desired period. It is more computationally intensive to check if certain values of $n$ produce periodic functions and so we were able to verify this conjecture only for $p$ up to 100.

\section{Haikus} 
Its periodic  \\
What came before is what's next \\
Its periodic \\ \\
No matter the $p$ \\
You can find a few good $n$\\
That is what we hope \\
